Canopy leaf area of a mature evergreen *Eucalyptus* woodland does not respond to elevated atmospheric [CO~2~] but tracks water availability

### 2015-08-07 manuscript.Rmd
R version auto-inserted; removed LAI anomaly methods description

## Data analysis
We used generalized additive models (GAM) for all analyses of LAI and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed LAI dynamics. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. Random effects were included to account for the grouping of data within FACE rings, except when we analysed average LAI across all rings (Fig. 7). For Fig. 7, we estimated the derivative of the fitted GAM, including confidence intervals for the derivative, following Simpson [-@Simpson2014; see also @Curtis2014].

\
For comparison between LAI from photos and PAR transmittance, we used linear regression. To test for effects of C~a~ treatment on $LAI_\tau$ and litter production, we used linear-mixed effects models with FACE ring as the random effect, and C~a~ treatment and date as factor variables.

\
To assess effects of eC~a~ on $LAI_\tau$, we accounted for pre-treatment differences in canopy structure, as follows. The average $LAI_\tau$ over the entire period was well correlated with basal area (BA) measured prior to the onset of the eC~a~ treatment across the FACE rings (Fig. 3). Because the EucFACE woodland is well established, the differences in BA are not due to C~a~ treatment effects, but instead represent natural variation within the woodland. We then calculated the deviation of actual $LAI_\tau$ from the fitted relationship with BA shown in Fig. 3, to yield a 'LAI anomaly'. This way, we effectively used BA as a covariate in the analysis to account for pre-treatment variation.

@@

>>>>

## Data analysis
We used generalized additive models (GAM) for all analyses of LAI and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R version `r r_version` [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed LAI dynamics. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. Random effects were included to account for the grouping of data within FACE rings, except when we analysed average LAI across all rings (Fig. 7). For Fig. 7, we estimated the derivative of the fitted GAM, including confidence intervals for the derivative, following Simpson [-@Simpson2014; see also @Curtis2014].

\
To test for effects of C~a~ treatment on LAI and litter production, we used linear-mixed effects models with FACE ring as the random effect, and C~a~ treatment and date as factor variables. The random effect accounts for the repeated measures nature of the data. For LAI, we further used pre-treatment basal area (BA) of each FACE ring as a covariate. Because the EucFACE woodland is well established, the differences in BA are not due to C~a~ treatment effects, but instead represent natural variation within the woodland.

**More details here;package, which anova for Table 1; and now we use BA as a covariate  **

@@

# Supporting information captions {.unnumbered}

**Figure S1**. Increase in gapfraction ($\tau_d$) and total litter production during the 4-month drought period in 2013, calculated for each ring separately.

**Figure S2**. Correlation between L estimated from diffuse canopy transmittance (LAI) and from canopy photos.

**Figure S3**. Comparison of canopy transmittance ($\tau$) during highly diffuse conditions (F~diff~ > 0.98), or using all PAR data without filtering.
